---
title: "Husky vs S&P"
author: "Mike"
date: "January 28, 2015"
output:
  html_document:
    keep_md: yes
---



#loading dependencies
```{r}

library(quantmod)


```

#Retriving Stock symbols Husky and S&P 500 from Yahoo, and using linecharts.

```{r}
getSymbols("^GSPC", src = "yahoo", from = "2014-09-03")
SP <- GSPC
lineChart(SP, theme = "white")
```

```{r}
getSymbols("HSE.TO", src = "yahoo", from = "2014-09-02")
Husky <- HSE.TO
lineChart(Husky, theme = "white")
          
```

#Checking dimensions and headers for each variable

```{r}
dim(SP)
head(SP)
dim(Husky)
head(Husky)

```

#Pulling out Adjusted stock from each symbol

```{r}
StockHusky <- Husky[,6]
SPAdj <- GSPC$GSPC.Adjusted
```

#Running a regression

```{r}
model <- lm(StockHusky ~ SPAdj)
summary(model)
anova(model)

```

#Running residual plots in time series, and parital plots 

```{r}

plot(residuals(model))
residuals(model)
plot(as.numeric(ts(residuals(model))) ~ as.numeric(ts(SPAdj)))


```

#Plot the Model in time series with Fitted vs Residuals

```{r}
plot(as.numeric(ts(fitted(model))), as.numeric(ts(residuals(model))), xlab="Fitted", ylab= "residuals")
```

#Running a qq form, histogram, and autocorrelation

```{r}

qqnorm(residuals(model), ylab="residuals")
qqline(residuals(model), col= "red", lwd = 2)
hist(residuals(model))
acf(residuals(model))
predict(model)

```
