---
title: "Week 1 assignment"
author: "Mike"
date: "January 14, 2015"
output:
  html_document:
    keep_md: yes
---
Michael Enquist

Reporting on Kinder Morgan and Embridge Pipelines


```{r}
# dependencies
library(car)
library(quantmod)
library(plyr)
library(dplyr)
```
Running linecharts for S&P 500, Kinder Morgan, and Embridge pipeline. From the beginning of January 2 2014 to January 21 2015

```{r}
#Checking out S&P 500 and running a line chart from January 2 2014
getSymbols("^GSPC", src = "yahoo", from = "2014-01-02" , to ="2015-02-10")
SP <- GSPC
lineChart(SP, theme = "white")
```

Using getSymbols from yahoo as a source, and retrieving S&P 500, Kinder Morgan and Embridge Pipeline.

```{r}
#Checking out Kinder Morgan and running a line chart from January 2 2014
getSymbols("KMI", src = "yahoo", from = "2014-01-02" , to ="2015-02-10")
Kinder <- KMI
lineChart(Kinder, theme = "white")
```



```{r}
#Checking out S&P 500 and running a line chart from January 2 2014
getSymbols("ENB.TO", src = "yahoo", from = "2014-01-02" , to ="2015-02-10")
Embridge <- ENB.TO
lineChart(Embridge, theme = "white")
```

Now checking out the adjusted stock for the two stock symbols and S&P 500 by using techniques that i learned from AQM. Changing the stocks and symbols into objects.

```{r}
# run structure for Kinder
str(Kinder)
```

The structure for Kinder Morgan is an "xts" object. The data is strucutred as Data: num [1:279, 1:6].

```{r}
# run dimensions for Kinder
dim(Kinder)
```

```{r}
# run head for Kinder
head(Kinder)
```

```{r}
# run Time for Kinder
time(Kinder)
```

Finding out the data object (Kinder) using the techniques from what i learned in the workshop - str(), dim(), head(), time(). The structure of KinderMorgan is a XTS object. It has 274 rows and 6 columns. The Columns have headers KMI.Open, KMI.High, KMI.Close, KMI.Volume, KMI.Adjusted. It shown the weekly returns from January 2 2014 to present.

```{r}
# change kinder morgan adjusted into object 
StockKMI <- KMI$KMI.Adjusted
# change S&P 500 adjeusted into object
SPAdj <- GSPC$GSPC.Adjusted
# change Embridge Pipeline adjusted into object 
StockENB <- ENB.TO$ENB.TO.Adjusted

#Kinder Morgan weekly stock returns
StockreturnKMI <- diff(StockKMI)/StockKMI[-length(StockKMI)][-1,]
KinderMorgan <- weeklyReturn(StockKMI)

#Embridge pipeline weekly returns
StockreturnENB <- diff(StockENB)/StockENB[-length(StockENB)][-1,]
EmbridgePipeline <- weeklyReturn(StockENB)


SPRet <- diff(SPAdj)/SPAdj[-length(SPAdj)][-1,]
SPreturn <- weeklyReturn(SPAdj)

```

I decided to call only the adjusted columns for each stock variable and S&P 500. I wanted the adjusted columns because it reveals the volatility returns from each week. I changed each stock adjusted into an object (ex. StockKMI <- KMI$KMI.Adjusted). 

 
```{r}

#Checkcing out the dimenasions for Kinder Morgan and S&P 500 so that each data objectare even out.
dim(KinderMorgan)
dim(SPreturn)
```

By checking out the dimensions 58 rows and 1 columns KinderMorgan and SPreturn. This means that i can run a regression for both of these variables.


```{r}
# Run a regression Kinder Morgan and S&P 500 and use as.vector as it is .xts flie
KinderMorgan <- data.frame(KinderMorgan = as.vector(KinderMorgan))
SPreturn <- data.frame(SPreturn = as.vector(SPreturn))
df <- data.frame(KinderMorgan, SPreturn)
# run  regression
model1 <- lm(KinderMorgan ~ SPreturn, data = df)
# Run Summary
summary(model1)

```

The return for Kinder Morgan and S&P 500 returns reveals p-value of 4.203e-05, and f-statistic of 19.76.  For every increase of the S&P 500, Kinder Morgan stock increases by 1.039 %. The significant codes reveals that the model is strong, however, the Residual standard error is 0.02832 , and the Adjusted R-squared is 0.2476. This means that my model can only explain a little about the volatility of Kinder Morgan stock from the past year to present. 

```{r}
#Running a plot for Kinder Morgan and S&P 500
plot(df)
abline(a = coef(model1)[1], b = coef(model1)[2])
 plot(df)
 abline(coef = coef(model1), col = "red", lwd = 2)
```

I created an abline through the data set for KinderMorgan and SPreturn. The line is arbitrary because it doesn't explain the line truly fits through the data set.

```{r}
#Running an anova table for Kinder Morgan and S&P 500
anova(model1)


```
The Anova table explains again the prediction model for Kinder Morgan and S&P 500 returns. The residuals are very small and only explain .044918 of the prediction model. Again this explain the S&P 500 returns are a weak explanation for Kinder Morgan Stock but reveals more information as the significance is '***'. So now i want to run dianostics because im not sure about this model.

```{r}
plot(as.numeric(ts(fitted(model1))), as.numeric(ts(residuals(model1))), xlab="Fitted", ylab= "residuals")
```

I am plotting the model1 to find a fitted vs residuals. I want to see if model1 can predict new data in the future. The x-axis is the "Fitted" and the y-axis is "residual".


```{r}
#Running a qq form, histogram, and autocorrelation
qqnorm(residuals(model1), ylab="residuals")
qqline(residuals(model1), col= "red", lwd = 2)
```

A Q–Q plot is used to compare the shapes of distributions, providing a graphical view of how properties such as location, scale, and skewness are similar or different in the two distributions. There are out-sized residuals, as indicated by the fat tails.  This casts some suspicion on the model, but might be explained by the excessive market volatility of from the past four months as the oil price as taken a hit. The plot is strongly s-shaped indicating that the daily percentage changes in the S&P 500 are not normally distributed.


```{r}
hist(residuals(model1))
```

Most of the data are on the left side of the histogram, but a few larger values are on the right, the data are said to be skewed to the right. The model1 has more negative returns than postive reutrns compared to the S & P 500. All the error should be close to zero however, the residuals very far from the zero. The normality of the errors is large and far from zero and it is not normally distributed. This distributon is skewed. This means that the stock is a risk. This model assumes a normal distribution of data when in fact data is skewed to the right of the mean.


```{r}
acf(residuals(model1))

```

ACF of the residuals from the  method applied to the S & P 500. The lack of correlation suggests the forecasts are not good. By looking at the ACF plot, we can do a more formal test for autocorrelation by considering a whole set of stock adjusted values as a group, rather than treat each one separately. When we look at the ACF plot to see if each spike is within the required limits with a small probability of giving a false positive. When enough of these tests are done, it is likely that at least one will give a false positive. The residuals should have some remaining autocorrelation, when in fact they do not, and it is shown by the chart.

 
```{r}
plot(residuals(model1))
abline(coef = coef(model1), col = "red", lwd = 2)
```


```{r}
crPlots(model1)
```

```{r}
predict(model1) 

```

```{r}
#Running a regression on Embridge and S&P 500 returns

dim(EmbridgePipeline)
dim(SPreturn)
```


```{r}

# Run a regression for EmbridgePipeline and S&P 500

EmbridgePipeline <- data.frame(EmbridgePipeline = as.vector(EmbridgePipeline))
SPreturn <- data.frame(SPreturn = as.vector(SPreturn))
df1 <- data.frame(EmbridgePipeline, SPreturn)
model2 <- lm(EmbridgePipeline ~ SPreturn, data = df1)
summary(model2)
```

The returns against Embridge pipeline and S&P 500 has shown p-value of .0002865, and f-statistic of 14.98, which is larger compared to Kinder Morgan.  For every increase of the S&P 500, Embridge pipeline stock increases by .891822 %. The significant codes reveals that the model is strong, however, the Residual standard error is 0.02791 , and the Adjusted R-squared is 0.1969. Yet again, my model can only explain a little about the volatility of Embridge pipeline stock just like Kinder Morgan stock.

```{r}
#Running a plot for Embridge Pipeline and S&P 500
plot(df1)
abline(coef = coef(model2), col = "red", lwd = 2)
   
```

An abline through the data set for Embridge pipeline and SPreturn. The line is arbitrary because it doesn't explain the line truly fits through the data set. However, it reveals some cluster points near the red line. This means that the events of Embridge and S&P 500 are not random and has some non- linearity.
```{r}
#running Anova for Embridge and S&P returns
anova(model2)

```

Running the Anova table shows the prediction model for Embride pipeline is strong as the significant codes is '***'. The residual plots are small as it is .043615



```{r}

#Now run a regression for Kinder Morgan and S&P 500
dim(EmbridgePipeline)
dim(KinderMorgan)
```

Now I decided to run a regression for Embridge Pipeline and Kinder Morgan. I want to see if my previous models are important to use for S&P 500 or not. 

```{r}
# Run a regression for Embridge Pipeline and Kinder Morgan

df2 <- data.frame(EmbridgePipeline, KinderMorgan)
model3 <- lm(EmbridgePipeline ~ KinderMorgan, data = df2)
summary(model3)

```

The regression against Embridge pipelines and Kinder Morgan is very weak. The p-value is .01009, and F-statistic is 7.092. The significant code is '*' which means the model can be changed or modified better.

```{r}
#Ploting returns on adjiusting closing prices for Embridge Pipeline against S&P 500
plot(df2)
abline(coef = coef(model3), col = "red", lwd = 2) 

```

The red line reveals a more cluster points compared to the other data sets. They are closer together which explains perhaps there is some correlation between returns for Kinder Morgan and Embridge Pipeline. However, there are some outliers which explains why these two stocks are not strongly correlated between each other. 

```{r}
#Running an anova table for Kinder Morgan and Embridge pipeline
anova(model3)


```
By analyzing the stock Kinder Morgan and Embridge pipeline. For every increase on for Kinder Morgan, there is 0.351520 percent increase in the Kinder Morgan Stock. The interaction between these two stocks are not correlated with S&P 500 returns. So which means this Model does not reveal on how the returns for these two stocks form Kinder Morgan and Embridge Pipelines are significant. 
This analysis has shown that the daily returns of the S&P 500 Index are poorly modeled 