---
title: "Oil&Gas Companies which impact the Canadian economy"
author: "Mike"
date: "November 8, 2014"
output:
  html_document:
    keep_md: yes
  pdf_document: default
---

I'm analyzing the major oil&drilling companies which impacts the Canadian Oil&Gas industry. I only included the companies that i felt impacts Alberta and northern BC, yet i know i can find more companies on yahoo.  The time period is around early 2000's to present. I worked in the Oil&Gas industry around early 2006 till 2010. Also, I included oil & natural gas prices from the past 10 years, and the US/Canadian Exchange rate as well. 
```{r}
library(knitr)
suppressPackageStartupMessages(library(quantmod))
```

```{r,results='asis'}
# Import data Patterson-Uti
year = "1900-01-01"
options("getSymbols.warning4.0"=FALSE)
Henderson.Land.Development <-getSymbols("0012.HK", src = "yahoo", from = "2000-01-01")
Henderson.Land.Development <- "0012.HK"
#Patterson-Uti is a land drilling company. They are a major player in US, they are small in Canada but they have enough capital to compete with Nabors and Precision Drilling.  
# Check data structure
kable(head(Henderson.Land.Development), format = "markdown")

# Plot data
lineChart(Henderson.Land.Development, theme = "white")

getSymbols("HSE.TO", src = "yahoo", from = "2014-01-02")
Husky <- HSE.TO
lineChart(Husky, theme = "white")


# load dependencies
library(quantmod)

# Import data
HAL <- getSymbols("0012.HK", src = "yahoo", from = "2008-01-01", to = "2009-11-28", auto.assign = FALSE)[,6]
Precison_Drilling <- HAL
str(Precison_Drilling)
# Check data structure
kable(head(Precison_Drilling), format = "markdown")
ggplot(Precison_Drilling, aes(y = '0012.HK.Adjusted', x = date)) + geom_point(size = 3, colour = "blue") + theme_bw()

Henderson.Land.Development <- as.data.frame(Precison_Drilling)
str(Henderson.Land.Development)
# Plot data
lineChart(HAL, theme = "white")

# load dependencies
library(quantmod)

# Import data
getSymbols("TOU.TO", src = "yahoo", from = "2000-01-01")
Tourmaline <- TOU.TO

# Check data structure
Tourmaline.tail <- tail(Tourmaline, n = 3650)

# Plot data
lineChart(Tourmaline, theme = "white")

# load dependencies
library(quantmod)

# Import data Canadian National Resources
getSymbols("CNQ.TO", src = "yahoo", from = "2000-01-01")
CNR <- CNQ.TO

# Check data structure
kable(head(CNR), format = "markdown")


# Plot data
chartSeries(CNR, theme = "white", type = "line")

# load dependencies
library(quantmod)

# Import data
getSymbols("ECA.TO", src = "yahoo", from = "2000-01-01")
Encana <- ECA.TO

# Check data structure
kable(head(Encana), format = "markdown")
Encana.tail <- tail(Encana, n = 3650)

# Plot data
lineChart(Encana, theme = "white")

# load dependencies
library(quantmod)

# Import data
getSymbols("BP", src = "yahoo", from = "2000-01-01")
BP <- BP

# Check data structure
kable(head(BP), format = "markdown")
BP.tail <- tail(BP, n = 3650)

# Plot data
lineChart(BP, theme = "white")

# load dependencies
library(quantmod)

# Import data
getSymbols("XOM", src = "yahoo")
Exxon <- XOM

# Check data structure
kable(head(Exxon), format = "markdown")
Exxon.tail <- tail(Exxon, n = 3650)

# Plot data
lineChart(Exxon, theme = "white")

# load dependencies
library(quantmod)

# Import data
getSymbols("ECA.TO", src = "yahoo", from = "2000-01-01")
Encana <- ECA.TO

# Check data structure
kable(head(Encana), format = "markdown")
Encana.tail <- tail(Encana, n = 3650)

# Plot data
lineChart(Encana, theme = "white")

# load dependencies
library(quantmod)

# Import data
Shell <- getSymbols("RDS-A", src = "yahoo",from = "2000-01-01", auto.assign = FALSE)


# Check data structure
kable(head(Shell), format = "markdown")

# Plot data
lineChart(Shell, theme = "white")

# load dependencies
library(quantmod)

# Import data
getSymbols("NBR", src = "yahoo", from = "2000-01-01")
Nabors <- NBR

# Check data structure
kable(head(Nabors), format = "markdown")

# Plot data
lineChart(Nabors, theme = "white")

# load dependencies
library(quantmod)

# Import data
getSymbols("HAL", src = "yahoo", from = "2000-01-01")
Halliburton <- HAL

# Check data structure
kable(head(Halliburton), format = "markdown")

# Plot data
lineChart(Halliburton, theme = "white")

# load dependencies
library(quantmod)

# Import data
getSymbols("SLB", src = "yahoo", from = "2000-01-01")
Schlumberger <- SLB

# Check data structure
kable(head(Schlumberger), format = "markdown")

# Plot data
lineChart(Schlumberger, theme = "white")

# load dependencies
library(quantmod)

# Import data
getSymbols("PSI.TO", src = "yahoo", from = "2000-01-01")
Pason <- PSI.TO

# Check data structure
kable(head(Pason), format = "markdown")

# Plot data
lineChart(Pason, theme = "white")

# load dependencies
library(quantmod)

# Import data 
getSymbols("SU", src = "yahoo", from = "2000-01-01")
Suncor <- SU

# Check data structure
kable(head(Suncor), format = "markdown")

# Plot data
chartSeries(Suncor, theme = "white", type = "line")

# load dependencies
library(quantmod)

# Import data 
getSymbols("CPG", src = "yahoo", from = "2000-01-01")
Crescent <- CPG

# Check data structure
kable(head(Crescent), format = "markdown")

# Plot data
chartSeries(Crescent, theme = "white", type = "line")

# load dependencies
library(quantmod)

# Import data 
getSymbols("COS.TO", src = "yahoo", from = "2000-01-01")
CANOILSANDS <- COS.TO

# Check data structure
kable(head(CANOILSANDS), format = "markdown")

# Plot data
chartSeries(CANOILSANDS, theme = "white", type = "line")

# load dependencies
library(quantmod)

# Import data 
getSymbols("CVE", src = "yahoo", from = "2000-01-01")
Cenovus <- CVE

# Check data structure
kable(head(Cenovus), format = "markdown")

# Plot data
chartSeries(Cenovus, theme = "white", type = "line")

# load dependencies
library(quantmod)

 # Import data
getSymbols("DCOILWTICO",src="FRED")
Oil_Prices <- DCOILWTICO
# Crude Oil Prices
# Check data structure
kable(head(Oil_Prices), format = "markdown")
DCOILWTICO.tail <- tail(Oil_Prices)

# Plot data
lineChart(Oil_Prices, theme = "white")

# load dependencies
library(quantmod)

 # Import data
getSymbols("GASPRICE",src="FRED")
NaturalGas <- GASPRICE
# Natural Gas prices till August 10 2013, I need to find a recent updated database, however it is still useful...
# Check data structure
kable(head(NaturalGas), format = "markdown")
GASPRICE.tail <- tail(NaturalGas, n = 175)

# Plot data
lineChart(GASPRICE.tail, theme = "white")

# load dependencies
library(quantmod)

 # Import data
getSymbols("EXCAUS",src="FRED")
CDNUS_RATE <- EXCAUS
# The Canaidan/US exchange rate
# Check data structure
kable(head(CDNUS_RATE), format = "markdown")
EXCAUS.tail <- tail(CDNUS_RATE, n = 190)

# Plot data
lineChart(EXCAUS.tail, theme = "white")

```

