---
title: "Coffee Commodity Prices Relationship between Brazil and Colombia"
author: "Michael Enquist"
date: "July 28, 2016"
output:
  pdf_document: default
  html_document:
    keep_md: yes
---

```{r,echo=FALSE, warning=FALSE, results='hide'}
library(knitr)
library(Quandl)
library(devtools)
library(plyr)
library(dplyr)
library(ggplot2)
```


```{r,echo=FALSE, warning=FALSE, results='hide'}
Quandl.api_key("c58SRnkLsYYrxJs6d74e")

### Brazil Exchange Rate
Columbiafx <- Quandl("CURRFX/COPUSD", start_date="2000-01-01", end_date="2015-12-31", collapse="monthly")

Brazilexch <- Quandl("BNP/USDBRL", start_date="2000-01-01", end_date="2015-12-31", collapse="monthly")

### Brazil Commodities and GDP


Brazil_Comm <- Quandl(c("ODA/PCOFFOTM_USD"), start_date="2000-01-01", end_date="2015-12-31",
                      collapse="monthly")

```

I want to prove that commodity price shocks affects developing countries such as Brazil and Colombia. I am using my own statistical analysis. If commodity prices affects each countries GDP and Currency Exchange then we can see this relationship through commodity chains. 

```{r,echo=FALSE, warning=FALSE, results='hide'}
names(Brazil_Comm) <- c("Date", "Coffee")
head(Brazil_Comm)


### I combined both the matrices using cbind. After renaming, we cleaned the data by removing the extra date column.
Brazil <- cbind(Brazilexch, Brazil_Comm)
Colombia <- cbind(Columbiafx,Brazil_Comm)
names(Brazil) <- c("Date", "FXBrazil", "Date1", "Coffee")
names(Colombia) <- c("Date", "Columbiafx", "Date1", "Coffee")
Brazil$Date1 <- NULL
Colombia$Date1 <- NULL
```


```{r,echo=FALSE, warning=FALSE}
### Brazil Full Regression  on its Currency and the Price of Coffww

Brazilregress  <- lm(FXBrazil ~ Coffee, data=Brazil)
summary(Brazilregress)
anova(Brazilregress)
plot(FXBrazil ~ Coffee, data=Brazil, xlab = "Coffee Price", ylab = "Brazilian Real")
abline(coef(Brazilregress), col = "blue")
### Colombia Full Regression 
```

By running a regression on FXBrazil ~ Coffee, the t-statistic has shown significance between Coffee prices and the Brazillian Real of 1.27e-05 ***. What is interesting here is that Brazilian Real has a negative relationship with the price of coffee and its own currency. I used historical data from 2000 to presetn using Quandl.  

```{r,echo=FALSE, warning=FALSE}


Colombiaregress  <- lm(Columbiafx ~ Coffee, data=Colombia)
summary(Colombiaregress)
anova(Colombiaregress)
plot(Columbiafx ~ Coffee, data=Colombia, xlab = "Coffee", ylab = "Colombian Peso")
abline(coef(Colombiaregress), col = "blue")
```

By running a regression on Columbiafx ~ Coffee, the t-statistic has shown significance between Coffee prices and the Colombian Peso of 2e-16 ***. What is interesting here is that Colombian Peso has a Positive relationship with the price of coffee and its own currency. 

```{r,echo=FALSE, warning=FALSE, results='hide'}

###### Creating a chart ##########################################################

Brazil_Comm <- Quandl(c("ODA/PCOFFOTM_USD"), start_date="1996-12-31", 
                      end_date="2014-01-31",collapse="annual")

Brazil.gdp <- Quandl("FRED/MKTGDPBRA646NWDB", 
                     start_date="1995-12-31", end_date="2014-12-31", 
                     collapse = "annual")

Colombia.gdp <- Quandl("FRED/MKTGDPCOA646NWDB", 
                       start_date="1995-12-31", end_date="2014-12-31", 
                       collapse = "annual")

Columbiafx <- Quandl("CURRFX/COPUSD", start_date="1996-12-31", end_date="2014-12-31", 
                     collapse="annual")

Brazilexch <- Quandl("FED/RXI_N_A_BZ", start_date="1996-12-31", end_date="2014-12-31", 
                     collapse="annual")

Brazil_Comm <- Quandl(c("ODA/PCOFFROB_USD"), start_date="1996-12-31", end_date="2014-12-31",
                      collapse="annual")

Brazil.Coffee.Prod <- Quandl("USDAFAS/COFFEE_PRODUCTION_BR", start_date="1996-12-31",
                      end_date="2014-12-31",collapse="annual")

Colombia.Coffee.Prod <- Quandl("USDAFAS/COFFEE_PRODUCTION_CO", start_date="1996-12-31",
                      end_date="2014-12-31")


```


```{r,echo=FALSE, warning=FALSE, results='hide'}

Brazil <- cbind(Brazilexch, Brazil_Comm)
Colombia <- cbind(Columbiafx,Brazil_Comm)
names(Brazil) <- c("Date", "FXBrazil", "Date1", "Coffee")
names(Colombia) <- c("Date", "Columbiafx", "Date1", "Coffee")
Brazil$Date1 <- NULL
Colombia$Date1 <- NULL
coffee <- cbind(Brazil,Colombia$Columbiafx)
names(coffee) <- c("Date", "FXBrazil", "Coffee", "Columbiafx")


GDP.join <- as.data.frame(inner_join(Brazil.gdp, Colombia.gdp, by = "DATE"))
names(GDP.join) <- c("Date","Brazil.gdp", "Colombia.gdp")
Prod.join <- as.data.frame(inner_join(Brazil.Coffee.Prod, Colombia.Coffee.Prod, by = "Year"))
names(Prod.join) <- c("Date","Brazil.Coffee.Prod", "Colombia.Coffee.Prod")
GDP.Prod.join <- as.data.frame(inner_join(GDP.join, Prod.join, by = "Date"))

```

```{r,echo=FALSE, warning=FALSE, results='hide'}
 
Merge.Metrics <- as.data.frame(inner_join(coffee, GDP.Prod.join, by = "Date"))


merged_rates <- Merge.Metrics %>%
  mutate(Rate.Brazil.FX = 100*((FXBrazil - lag(FXBrazil)))/FXBrazil) %>%
  mutate(Rate.Columbia.FX = 100*((Columbiafx - lag(Columbiafx)))/Columbiafx) %>%
  mutate(Rate.Coffee = 100*((Coffee - lag(Coffee)))/Coffee) %>% 
  mutate(Rate.Brazil.gdp = 100*((Brazil.gdp - lag(Brazil.gdp)))/Brazil.gdp) %>%
  mutate(Rate.Colombia.gdp = 100*((Colombia.gdp - lag(Colombia.gdp)))/Colombia.gdp) %>% 
  mutate(Rate.Brazil.Coffee.Prod = 100*((Brazil.Coffee.Prod - lag(Brazil.Coffee.Prod)))/Brazil.Coffee.Prod) %>% 
  mutate(Rate.Colombia.Coffee.Prod = 100*((Colombia.Coffee.Prod - lag(Colombia.Coffee.Prod)))/Colombia.Coffee.Prod) 


```

```{r,echo=FALSE, warning=FALSE, results='hide'}
#labels=c("Brazilian Real","Colombian Peso","Commodity Price of Coffee")

Coffee_Market <- ggplot(merged_rates, aes(x =Date)) + 
  geom_line(aes(y = Rate.Brazil.FX, colour = "Brazilian Real")) + 
  geom_line(aes(y = Rate.Coffee, colour = "Commodity Price of Coffee")) +
  theme(plot.title = element_text(size=15, face="bold", hjust=2)) + theme_bw() +
  scale_fill_continuous(name="Percentage.Change")  + ylab("Percentage.Change") + xlab("Years 1996 to 2015")
Coffee_Market 
```


What i did here is created a rate of change function on Coffee, and the Currencies on Brazil and Colombia from 1996 to present. The statisitical analysis can explain the positive adn negative relationship between currencies and coffee price. 



```{r,echo=FALSE, warning=FALSE, results='hide'}

####### GDP


Coffee_GDP <- ggplot(merged_rates, aes(x =Date)) + 
  geom_line(aes(y = Rate.Brazil.gdp, colour = "Brazilian GDP")) + 
  geom_line(aes(y = Rate.Coffee, colour = "Commodity Price of Coffee")) +
  theme(plot.title = element_text(size=15, face="bold", hjust=1)) + theme_bw() +
  scale_fill_continuous(name="Percentage.Change") + ylab("Percentage.Change") +xlab("Years 1996 to 2015")
Coffee_GDP
```


What i did here is created a rate of change function on Coffee, and the GDP on Brazi. This chart explains each economy is affected by coffee prices and their GDP. 

```{r,echo=FALSE, warning=FALSE, results='hide'}
Coffee_Prod <- ggplot(merged_rates, aes(x =Date)) + 
  geom_line(aes(y = Rate.Brazil.Coffee.Prod, colour = "Brazilian Coffee Production")) + 
  geom_line(aes(y = Rate.Coffee, colour = "Commodity Price of Coffee")) +
  theme(plot.title = element_text(size=15, face="bold", hjust=2)) + theme_bw() +
  scale_fill_continuous(name="Percentage.Change") + ylab("Percentage.Change")+ xlab("Years 1996 to 2015")
Coffee_Prod
```
By looking at this graph, the production of coffee from each country and the price of coffee is quite volitile. Brazil production quite considerably from the past 20 years. 

```{r,echo=FALSE, warning=FALSE}
### Brazil Full Regression  on its GDP and the Price of Coffee

Brazilregress  <- lm(Brazil.gdp ~ Coffee, data=Merge.Metrics)
summary(Brazilregress)
anova(Brazilregress)
plot(Brazil.gdp ~ Coffee, data=Merge.Metrics, xlab = "Coffee Price", ylab = "Brazil GDP")
abline(coef(Brazilregress), col = "blue")
### Colombia Full Regression 
```


```{r,echo=FALSE, warning=FALSE}
### Brazil Full Regression  on its Coffee Production and the Price of Coffee

Brazilregress  <- lm(Brazil.Coffee.Prod ~ Coffee, data=Merge.Metrics)
summary(Brazilregress)
anova(Brazilregress)
plot(Brazil.Coffee.Prod ~ Coffee, data=Merge.Metrics, xlab =  "Coffee Price", ylab = "Brazilian Coffee Produciton")
abline(coef(Brazilregress), col = "blue")
### Colombia Full Regression 
```

```{r,echo=FALSE, warning=FALSE}
### Brazil Full Regression  on its Currency and the Price of Coffww

Brazilregress  <- lm(FXBrazil ~ Coffee, data=Merge.Metrics)
summary(Brazilregress)
anova(Brazilregress)
plot(FXBrazil ~ Coffee, data=Merge.Metrics, xlab = "Coffee Price", ylab = "Brazilian Real")
abline(coef(Brazilregress), col = "blue")
### Colombia Full Regression 








